---
title: "Part III: Plotting, supplementation, and export of stats dataset"
author: "Mike Vlah"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "documentation") })
---

# Description

<font size="5">Summary of major workflow components</font> 

This workflow is part of a series that describes the preparation of data used in [Bernhardt et al. (2022)](https://doi.org/10.1073/pnas.2121976119). The creation of this dataset consists of several sub workflows as described below.

* **Part I: Standardized dataset creation:** Part I is the preparation of standardized datasets of stream metabolism from the [StreamPULSE data portal](https://data.streampulse.org/) and terrestrial ecosystem fluxes from the [FLUXNET2015 dataset](https://fluxnet.org/data/fluxnet2015-dataset/). Everything generated in Part I is included in `output_data`, so you may start with Part II.

* **Part II. Filtering, gap-filling, and calculating metrics:** In Part II the standardized dataset of lotic and terrestrial metabolism data was filtered down to the subset of sites used in Bernhardt et al. (2022). After filtering, additional descriptive metrics were calculated to use in analysis.

* **Part III. Plotting and export of stats dataset:** Datasets from part II are minimally subsetted and recast for plotting convenience. Figures 1-6 are generated.

* **Part IV. Structural equation modeling:** Code used for structural equation model (SEM) analysis on annual metabolism dataset. We used an observed variables model to estimates the effect of light (PAR reaching the stream surface) and hydrologic variability (skewness of daily discharge) on annual river GPP.  

# Setup

See code for details

```{r setup, message = FALSE}
library(plyr)
library(tidyverse)
library(RColorBrewer)
library(plotrix)
library(here)
```

```{r setup2, message = FALSE}

dir.create(here('figures'),
           showWarnings = FALSE)

#choose colors for plot categories
fnetcolor = 'sienna4'
spcolor = 'cadetblue4'
gppcolor = 'forestgreen'
ercolor = 'sienna'

source(here('plot_helpers.R'))
```

# Load site data from Part II

Columns Stream_PAR_sum, Disch_ar1, MOD_ann_NPP will be used in plots.

```{r load site data, message = FALSE}

site_data_1 = readRDS(here('output_data', 'lotic_site_info_filtered.rds')) %>%
    as_tibble() %>%
    select(sitecode = Site_ID,
           Stream_PAR_sum, Disch_ar1, Disch_skew, MOD_ann_NPP, nyears, ndays, coverage)

```

# Load and prepare fluxnet data

We're beginning from two previously generated datasets here. The first is `fluxnet_filtered_metabolism.rds`, a collection of data.frames for each terrestrial site, organized in a list. We'll cast that to one big tibble, select and rename a few columns, calculate NEP from GPP and ER, and determine a mean value for each of those metrics for each DOY, across years. The resulting dataset (summarized by site and DOY) is the basis for the "lips plots" that show median GPP, ER, and NEP (and GPP and ER ranges) over the course of an average year. For every other plot, we use a tibble that summarizes terrestrial data by site, called `fnet_site`.

```{r load site data fnet, message = FALSE}

restricted_use_sites = c("RU-Sam", "RU-SkP", "RU-Tks", "RU-Vrk", "SE-St1", "ZA-Kru")

#summarize by site, starting with a dataset that's summarized by site-year
fnet_ann = readRDS(here('output_data', 'fluxnet_filtered_metabolism.rds'))
fnet_names = names(fnet_ann)

for(i in 1:length(fnet_ann)){
    fnet_ann[[i]]$sitecode = fnet_names[i]
}

fnet_ann = fnet_ann[! fnet_names %in% restricted_use_sites]

fnet_full = Reduce(bind_rows, fnet_ann) %>%
    as_tibble() %>%
    filter(sitecode != 'CZ-BK2')

fnet_site = readRDS(here('output_data', 'fluxnet_site_info_filtered.rds')) %>%
    as_tibble() %>%
    select(sitecode = Site_ID,
           GPP_site_mean = ann_GPP,
           ER_site_mean = ann_ER) %>%
    filter(! is.na(GPP_site_mean), #CZ-BK2 has NaN
           ! is.na(ER_site_mean)) %>%
    mutate(NEP_site_mean = GPP_site_mean + ER_site_mean) %>%
    arrange(sitecode)

#summarize by DOY for lips plots
fnet_lips = fnet_full %>%
    select(sitecode, GPP, ER, DOY, Year) %>%
    group_by(sitecode, DOY) %>%
    summarize(GPP = mean(GPP, na.rm=TRUE), #average metab by day across years
              ER = mean(ER, na.rm=TRUE)) %>%
    ungroup() %>%
    arrange(sitecode, DOY) %>%
    rename(GPP_C_filled = GPP, ER_C_filled = ER) %>%
    mutate(NEP_C_filled = GPP_C_filled + ER_C_filled)
```

# Load and prepare streampulse data

Same procedures as for FLUXNET (terrestrial) data, here applied to StreamPULSE (aquatic) data.

```{r load site data sp, message = FALSE}
sp_list = readRDS(here('output_data', 'lotic_gap_filled.rds'))

#summarize by site
sp_names = names(sp_list)

for(i in 1:length(sp_list)){
    sp_list[[i]]$sitecode = sp_names[i]
}

sp_full = Reduce(bind_rows, sp_list) %>%
    as_tibble() %>%
    select(sitecode, Date, Year, DOY, GPP_C, ER_C, GPP_C_filled, ER_C_filled) %>%
    arrange(sitecode, Date)

sp_site = sp_full %>%
    group_by(sitecode, Year) %>%
    summarize(GPP_ann_sum = sum(GPP_C_filled, na.rm = TRUE),
              ER_ann_sum = sum(ER_C_filled, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(sitecode) %>%
    summarize(GPP_site_mean = mean(GPP_ann_sum, na.rm = TRUE),
              ER_site_mean = mean(ER_ann_sum, na.rm = TRUE)) %>%
    ungroup() %>%
    arrange(sitecode) %>%
    mutate(NEP_site_mean = GPP_site_mean + ER_site_mean) %>%
    left_join(site_data_1, by = 'sitecode')

#summarize by DOY for lips plots
sp_lips = sp_full %>%
    select(sitecode, Year, DOY, GPP_C_filled, ER_C_filled) %>%
    group_by(sitecode, DOY) %>%
    summarize(GPP_C_filled = mean(GPP_C_filled, na.rm=TRUE), #average metab by day across years
              ER_C_filled = mean(ER_C_filled, na.rm=TRUE)) %>%
    ungroup() %>%
    arrange(sitecode, DOY) %>%
    mutate(NEP_C_filled = GPP_C_filled + ER_C_filled) %>%
    left_join(site_data_1, by = 'sitecode') #include site data
```

# Split sites by coverage

For some plots, we use different markers to indicate sites with high coverage for metabolism estimates (>= 80% of days across years) and those with low coverage (< 80% of days). Here we create vectors of sitecodes that fit each of these criteria, and boolean vectors representing the same, to be used for subsetting later.

```{r split by coverage, message = FALSE}

coverage_sp = site_data_1 %>%
    select(sitecode, ndays, nyears, coverage)

coverage_fnet = readRDS(here('output_data', 'fluxnet_site_info_filtered.rds')) %>%
    as_tibble() %>%
    select(sitecode = Site_ID,
           nyears, ndays, coverage) %>%
    filter(sitecode != 'CZ-BK2')

sp_high_cov_sites = coverage_sp %>%
    filter(coverage >= 0.8) %>%
    pull(sitecode)
sp_high_cov_bool = sp_site$sitecode %in% sp_high_cov_sites

fnet_high_cov_sites = coverage_fnet %>%
    filter(coverage >= 0.8) %>%
    pull(sitecode)
fnet_high_cov_bool = fnet_site$sitecode %in% fnet_high_cov_sites
```

# Figure 1: GPP-ER biplot and dist plots

See code/paper for details

```{r fig 1, include = FALSE}
axis_cex = 2.4 #applies to labels and tick values

jpeg(width=10, height=10, units='in', res=300, quality=100, type='cairo',
     filename=here('figures', 'gpp_er_biplot_cumulAnnual.jpeg'))
# par(mar=c(4.5, 4.5, 2, 2))
par(mar=c(6.5, 6.5, 2, 2))

log_gpp_fnet = log(fnet_site$GPP_site_mean)
log_er_fnet = log(fnet_site$ER_site_mean * -1) * -1
log_gpp_sp = log(sp_site$GPP_site_mean)
log_er_sp = log(sp_site$ER_site_mean * -1) * -1

gpptck = c(1:9 %o% 10^(0:3))
ertck = rev(c(1:9 %o% 10^(0:3)))

# plot(log_gpp_fnet,
#      log_er_fnet, col=alpha(fnetcolor, alpha=0.5),
plot(log_gpp_fnet[fnet_high_cov_bool],
     log_er_fnet[fnet_high_cov_bool], col=alpha(fnetcolor, alpha=0.5),
     xlab='', ylab='', bg=alpha(fnetcolor, alpha=0.5),
     cex=1.5, cex.lab=axis_cex, cex.axis=axis_cex, ylim=-log(c(5000, 5)),
     pch=21, yaxt='n', xaxt='n', xlim=log(c(20, 5000)), lwd=2)
mtext(expression(paste("Annual GPP (g C"~"m"^"-2"*" y"^"-1"*')')),
      1, line=5, cex=axis_cex)
mtext(expression(paste("Annual ER (g C"~"m"^"-2"*" y"^"-1"*')')),
      2, line=3.5, cex=axis_cex)
# points(log_gpp_sp, log_er_sp, lwd=2,
#        col=alpha(spcolor, alpha=0.5), cex=1.5, pch=21, bg=alpha(spcolor, alpha=0.5))
points(log_gpp_sp[! sp_high_cov_bool], log_er_sp[! sp_high_cov_bool], lwd=2,
       col=alpha(spcolor, alpha=0.5), cex=1.5, pch=21, bg='transparent')
points(log_gpp_fnet[! fnet_high_cov_bool], log_er_fnet[! fnet_high_cov_bool],
       col=alpha(fnetcolor, alpha=0.5), cex=1.5, pch=21, bg='transparent', lwd=2)
points(log_gpp_sp[sp_high_cov_bool], log_er_sp[sp_high_cov_bool], lwd=2,
       col=alpha(spcolor, alpha=0.5), cex=1.5, pch=21, bg=alpha(spcolor, alpha=0.5))
legend('topright', legend=c('FLUXNET', 'StreamPULSE'), pch=21, bty='n', pt.cex=1.5,
       col=c(alpha(fnetcolor, alpha=0.5), alpha(spcolor, alpha=0.5)), pt.lwd=2,
       pt.bg=c(alpha(fnetcolor, alpha=0.5), alpha(spcolor, alpha=0.5)), x.intersp=2)
legend('topright', legend=c('FLUXNET', 'StreamPULSE'), pch=21, bty='n', pt.cex=1.5,
       col=c(alpha(fnetcolor, alpha=0.5), alpha(spcolor, alpha=0.5)),
       pt.bg='transparent', pt.lwd=2)
all_gpp = c(fnet_site$GPP_site_mean, sp_site$GPP_site_mean)
all_gpp[all_gpp <= 0] = NA
gpprng = range(all_gpp, na.rm=TRUE)
all_er = c(fnet_site$ER_site_mean, sp_site$ER_site_mean)
all_er[all_er >= 0] = NA
errng = range(all_er, na.rm=TRUE)

gpptck_log = log(gpptck)
axis(1, at=gpptck_log, labels=rep('', length(gpptck_log)), cex.axis=axis_cex, padj=0.4, tcl=-0.4)
axis(1, at=log(c(20, 100, 1000, 5000)), labels=c(20, 100, 1000, 5000),
     cex.axis=axis_cex, padj=0.4, lwd.ticks=2, tcl=-.6)
ertck_log = log(ertck) * -1
axis(2, at=ertck_log, labels=rep('', length(ertck_log)), cex.axis=axis_cex, padj=0.4, tcl=-0.4)
axis(2, at=log(c(5000, 1000, 100, 10, 1)) * -1, labels=c(5000, 1000, 100, 10, 1) * -1,
     cex.axis=axis_cex, padj=0.2, lwd.ticks=2, tcl=-.6)

abline(a=0, b=-1, lty=2)

dev.off()

#--- distplots

jpeg(width=8, height=8, units='in', res=300, quality=100, type='cairo',
     filename=here('figures', 'gpp_er_distplots.jpeg'))

par(mfrow=c(2, 1), mar=c(2,1,1,1), oma=c(0, 0, 0, 0))

#GPP
dens = density(na.omit(log_gpp_fnet))
# dens = density(na.omit(log_gpp_fnet[fnet_high_cov_bool]))
gpp_dens_fnet = tibble(x=dens$x, y=dens$y)
dens = density(na.omit(log_gpp_sp))
# dens = density(na.omit(log_gpp_sp[sp_high_cov_bool]))
gpp_dens_sp = tibble(x=dens$x, y=dens$y)

plot(gpp_dens_fnet$x, gpp_dens_fnet$y, type='n', ann=FALSE, xaxt='n', yaxt='n',
     bty='n', xlim=c(-2,9.5))
axis(1, at=gpptck_log, labels=gpptck, padj=-1.3, tick = TRUE, line=-0.2, tcl=-0.2)
mtext('GPP', 1, line=1)
polygon(x=c(gpp_dens_fnet$x, rev(gpp_dens_fnet$x)),
        y=c(gpp_dens_fnet$y, rep(0, nrow(gpp_dens_fnet))),
        col=alpha(fnetcolor, alpha=0.7),
        border=alpha(fnetcolor, alpha=0.7))
polygon(x=c(gpp_dens_sp$x, rev(gpp_dens_sp$x)),
        y=c(gpp_dens_sp$y, rep(0, nrow(gpp_dens_sp))),
        col=alpha(spcolor, alpha=0.7),
        border=alpha(spcolor, alpha=0.7))

#ER
dens = density(na.omit(log_er_fnet))
# dens = density(na.omit(log_er_fnet[fnet_high_cov_bool]))
er_dens_fnet = tibble(x=dens$x, y=dens$y)
dens = density(na.omit(log_er_sp))
# dens = density(na.omit(log_er_sp[sp_high_cov_bool]))
er_dens_sp = tibble(x=dens$x, y=dens$y)

plot(er_dens_fnet$x, er_dens_fnet$y, type='n', ann=FALSE, xaxt='n', yaxt='n',
     bty='n', xlim=c(2, -9.5))
axis(1, at=ertck_log, labels=ertck, padj=-1.3, tick = TRUE, line=-0.2, tcl=-0.2)
mtext('ER', 1, line=1)
polygon(x=c(rev(er_dens_fnet$x), er_dens_fnet$x),
        y=c(rev(er_dens_fnet$y), rep(0, nrow(er_dens_fnet))),
        col=alpha(fnetcolor, alpha=0.7),
        border=alpha(fnetcolor, alpha=0.7))
polygon(x=c(er_dens_sp$x, rev(er_dens_sp$x)),
        y=c(rev(er_dens_sp$y), rep(0, nrow(er_dens_sp))),
        col=alpha(spcolor, alpha=0.7),
        border=alpha(spcolor, alpha=0.7))

dev.off()

#NEP
nep_fnet = fnet_site$NEP_site_mean
nep_sp = sp_site$NEP_site_mean

jpeg(width=8, height=8, units='in', res=300, quality=100, type='cairo',
     filename=here('figures', 'nep_distplot.jpeg'))

par(mfrow=c(2, 1), mar=c(2,1,1,1), oma=c(0, 0, 0, 0))

dens = density(na.omit(nep_fnet))
# dens = density(na.omit(nep_fnet[fnet_high_cov_bool]))
nep_dens_fnet = tibble(x=dens$x, y=dens$y)
dens = density(na.omit(nep_sp))
# dens = density(na.omit(nep_sp[sp_high_cov_bool]))
nep_dens_sp = tibble(x=dens$x, y=dens$y)

plot(nep_dens_sp$x, nep_dens_sp$y, type='n', ann=FALSE, xaxt='n', yaxt='n',
     bty='n', xlim=c(-1723, 2116))
axis(1, padj=-1.3, tick = TRUE, line=-0.2, tcl=-0.2)
mtext('NEP', 1, line=1)
polygon(x=c(rev(nep_dens_fnet$x), nep_dens_fnet$x),
        y=c(rev(nep_dens_fnet$y), rep(0, nrow(nep_dens_fnet))),
        col=alpha(fnetcolor, alpha=0.7),
        border=alpha(fnetcolor, alpha=0.7))
polygon(x=c(nep_dens_sp$x, rev(nep_dens_sp$x)),
        y=c(rev(nep_dens_sp$y), rep(0, nrow(nep_dens_sp))),
        col=alpha(spcolor, alpha=0.7),
        border=alpha(spcolor, alpha=0.7))

dev.off()
```

# Figures 2 and 3: lips plots and probability densities

See code/paper for details

```{r Figs 2 and 3, include = FALSE}
dir.create(here('figures' ,'lips'), showWarnings = FALSE)
dir.create(here('figures', 'lips', 'lips_cross'), showWarnings = FALSE)
dir.create(here('figures', 'probdens'), showWarnings = FALSE)

axis_cex = 2.4 #applies to labels and tick values

lips_plot = function(quant_filt=NULL, outfile, use_fnet=FALSE, ylims=NULL,
                     wee=FALSE){

    dset = if(use_fnet) fnet_lips else sp_lips
    dset_site = if(use_fnet) fnet_site else sp_site

    var_quant_filt = NULL
    if(! is.null(quant_filt)){
        
        filter_expr <- c()
        for(i in seq_along(quant_filt)){

            quant_comp = strsplit(quant_filt[i], ' ')[[1]]
            qf = quantile(dset_site[, quant_comp[1], drop=TRUE], na.rm=TRUE,
                          probs = as.numeric(quant_comp[3]))

            filter_expr = c(filter_expr,
                            paste(quant_comp[1], quant_comp[2], qf))
        }

        if(length(filter_expr) > 1){
            filter_expr = paste(filter_expr, collapse = ' & ')
        }

        filt_sites = dset_site %>%
            filter_(filter_expr) %>%
            pull(sitecode)
        dset = filter(dset, sitecode %in% filt_sites)
    }

    nsites_included = length(unique(dset$sitecode))

    smry = dset %>%
        select(DOY, GPP_C_filled, ER_C_filled, NEP_C_filled) %>%
        group_by(DOY) %>%
        summarize_all(list(median=~median(., na.rm=TRUE),
                           quant25=~quantile(., na.rm=TRUE)[2],
                           quant75=~quantile(., na.rm=TRUE)[4])) %>%
        ungroup()

    jpeg(width=11, height=11, units='in', filename=outfile, type='cairo',
         res=300, quality=100)
    # pdf(file=outfile, width=10, height=10)

    if(! is.null(ylims)){

        # plot_dims = top_dims = dev.size('in')
        # top_dims[2] = plot_dims[2] * abs(ylims[2]) / sum(abs(ylims))

        gpplim = c(0, ylims[2])
        erlim = c(ylims[1], 0)

    } else {

        # top_dims = dev.size('in')

        gpplim=c(0, 5)
        erlim=c(-5, 0)

        if(use_fnet){
            gpplim[2] = 11
            erlim[1] = -7
        }
    }

    if(wee){
        axis_cex_ = axis_cex * 2
        line_ = -1.5
        par(mfrow=c(2, 1), oma=c(1, 1, 0, 0), mar=c(0, 5, 3, 1), lend=2)
    } else {
        axis_cex_ = axis_cex
        line_ = -2
        par(mfrow=c(2, 1), oma=c(1, 1, 0, 0), mar=c(0, 5, 3, 1), lend=2)
    }
    # fin=top_dims, new=TRUE)

    plot(smry$DOY, smry$GPP_C_filled_median, ylab='', yaxs='i', type='l',
         bty='n', lwd=4, xlab='', ylim=gpplim, xaxs='i', xaxt='n', yaxt='n',
         col='gray30')
    polygon(x=c(smry$DOY, rev(smry$DOY)),
            y=c(smry$GPP_C_filled_quant25, rev(smry$GPP_C_filled_quant75)),
            border=NA, col=alpha('forestgreen', alpha=0.6))
    axislocs = if(max(ylims) >= 10) seq(0, 10, 2) else c(0, 2, 4)
    axis(2, las=0, line=0, xpd=NA, tck=-.02, labels=FALSE,
         at=axislocs, cex.axis=axis_cex_, tcl=-0.3)
    # at=round(seq(0, gpplim[2], length.out=5), 1))
    axis(2, las=0, line=-0.5, xpd=NA, tcl=0, col='transparent',
         at=axislocs, cex.axis=axis_cex_)
    # at=round(seq(0, gpplim[2], length.out=5), 1))
    abline(h=0, lty=2, lwd=2, col='gray60')
    medsums = round(colSums(select(smry, contains('median'))), 1)

    if(! wee){
        mtext(expression(paste(bold("g C") ~ bold("m") ^ bold("-2") *
                                   bold(" d") ^ bold("-1"))), side=2,
              line=line_, outer=TRUE, cex=axis_cex_)
    }

    # if(filter_label){
    #     legend('topright', title='Filters', bty='n', title.col='gray30',
    #            lty=1, seg.len=0.2, lwd=2, legend=c(..., var_quant_filt))
    # }

    legend('right', title='Cumulative\nMedian Sums', bty='n',
           legend=c(paste('GPP:', medsums[1]), paste('ER:', medsums[2]),
                    paste('NEP:', medsums[3])), title.col='gray30')
    legend('left', paste('Sites included:', nsites_included), bty='n')

    # if(! is.null(ylims)){
    #     plot_dims = bottom_dims = dev.size('in')
    #     bottom_dims[2] = plot_dims[2] * abs(ylims[1]) / sum(abs(ylims))
    # } else {
    #     bottom_dims = dev.size('in')
    # }

    par(mar=c(4, 5, 0, 1))#, fin=bottom_dims)
    if(wee){
        padj_ = 0.8
        DOY1 = ''
    } else {
        padj_ = 0.5
        DOY1 = 1
    }

    plot(smry$DOY, smry$ER_C_filled_median, ylab='', yaxs='i', type='l',
         bty='n', lwd=4, xlab='', ylim=erlim, xaxs='i', xaxt='n', yaxt='n')
    polygon(x=c(smry$DOY, rev(smry$DOY)),
            y=c(smry$ER_C_filled_quant25, rev(smry$ER_C_filled_quant75)),
            border=NA, col=alpha('sienna', alpha=0.6))
    if(max(ylims) >= 10){
        axislocs = seq(0, -10, -2)
        axisextra = -11
    } else {
        axislocs = c(0, -2, -4)
        axisextra = NULL
    }
    axis(2, las=0, line=0, xpd=NA, tck=-.02, labels=FALSE,
         # at=round(seq(0, erlim[1], length.out=5), 1))
         at=c(axislocs, axisextra), cex.axis=axis_cex_, tcl=-0.3)
    axis(2, las=0, line=-0.5, xpd=NA, tcl=0, col='transparent',
         # at=round(seq(0, erlim[1], length.out=5), 1))
         at=axislocs, cex.axis=axis_cex_)
    axis(1, line=0, tck=-.02, labels=FALSE, at=c(1, seq(60, max(smry$DOY), 60)),
         cex.axis=axis_cex_, tcl=-0.3)
    axis(1, line=-0.5, tcl=0, col='transparent', at=c(1, seq(60, max(smry$DOY), 60)),
         cex.axis=axis_cex_, padj = padj_, tcl=-0.3,
         labels = c(DOY1, seq(60, max(smry$DOY), 60)))
    lines(smry$DOY, smry$NEP_C_filled_median, col='black', lwd=4, xpd=NA, lend=1)

    if(! wee){
        mtext('DOY', side=1, line=3.5, font=2, cex=axis_cex_)
    }
    dev.off()
}

pdf_plot = function(var, outfile){

    jpeg(width=7, height=4, units='in', filename=outfile, type='cairo',
         res=300, quality=100)
    # pdf(outfile, height=4, width=7)

    vv = na.omit(sort(sp_site[[var]]))
    dens = density(vv)
    vq = quantile(vv, probs=c(0.25, 0.75))
    densdf = tibble(x=dens$x, y=dens$y)
    dens25 = dens75 = densdf
    dens25 = dens25[densdf$x <= vq[1], ]
    dens75 = dens75[densdf$x >= vq[2], ]
    plot(densdf$x, densdf$y, type='l', xlab='width', ylab='density', bty='l',
         col='gray50', lwd=2)
    polygon(x=c(dens25$x, rev(dens25$x)),
            y=c(dens25$y, rep(0, nrow(dens25))), col='gray50', border='gray50')
    polygon(x=c(dens75$x, rev(dens75$x)),
            y=c(dens75$y, rep(0, nrow(dens75))), col='gray50', border='gray50')

    dev.off()
}

#overall
lips_ylim = c(-11, 11)
lips_plot(quant_filt=NULL, outfile=here('figures', 'lips', 'lips_overall_sp.jpeg'),
          ylims=lips_ylim)
lips_plot(quant_filt=NULL, outfile=here('figures', 'lips', 'lips_overall_fnet.jpeg'),
          use_fnet=TRUE, ylims=lips_ylim)

pdf_plot('Disch_ar1', here('figures', 'probdens', 'probdens_Qar1.jpeg'))
pdf_plot('MOD_ann_NPP', here('figures', 'probdens', 'probdens_MODIS.jpeg'))
pdf_plot('Stream_PAR_sum', here('figures', 'probdens', 'probdens_PAR.jpeg'))

#subsets
lips_ylim = c(-5, 5)
lips_plot(quant_filt='Disch_skew > 0.75', outfile=here('figures', 'lips', 'lips_Qskew_75.jpeg'),
          ylims=lips_ylim, wee=TRUE)
lips_plot(quant_filt='Disch_skew < 0.25', outfile=here('figures', 'lips', 'lips_Qskew_25.jpeg'),
          ylims=lips_ylim, wee=TRUE)
lips_plot(quant_filt='MOD_ann_NPP > 0.75', outfile=here('figures', 'lips', 'lips_MODIS_75.jpeg'),
          ylims=lips_ylim, wee=TRUE)
lips_plot(quant_filt='MOD_ann_NPP < 0.25', outfile=here('figures', 'lips', 'lips_MODIS_25.jpeg'),
          ylims=lips_ylim, wee=TRUE)
lips_plot(quant_filt='Stream_PAR_sum > 0.75', outfile=here('figures', 'lips', 'lips_PAR_75.jpeg'),
          ylims=lips_ylim, wee=TRUE)
lips_plot(quant_filt='Stream_PAR_sum < 0.25', outfile=here('figures', 'lips', 'lips_PAR_25.jpeg'),
          ylims=lips_ylim, wee=TRUE)

#cross-subsets
lips_plot(quant_filt = c('Stream_PAR_sum < 0.25', 'Disch_skew < 0.25'),
          outfile = 'figures/lips/lips_cross/lips_PAR25-Qskew25.jpeg',
          ylims = lips_ylim,
          wee = TRUE)
lips_plot(quant_filt = c('Stream_PAR_sum < 0.25', 'Disch_skew > 0.75'),
          outfile = 'figures/lips/lips_cross/lips_PAR25-Qskew75.jpeg',
          ylims = lips_ylim,
          wee = TRUE)
lips_plot(quant_filt = c('Stream_PAR_sum > 0.75', 'Disch_skew > 0.75'),
          outfile = 'figures/lips/lips_cross/lips_PAR75-Qskew75.jpeg',
          ylims = lips_ylim,
          wee = TRUE)
lips_plot(quant_filt = c('Stream_PAR_sum > 0.75', 'Disch_skew < 0.25'),
          outfile = 'figures/lips/lips_cross/lips_PAR75-Qskew25.jpeg',
          ylims = c(-5, 5.6),
          wee = TRUE)
```

# Figure 4: bubble plots

See code/paper for details

```{r fig 4, include = FALSE}
dir.create(here('figures', 'bubble_plots'), showWarnings = FALSE)

bubble_plot = function(xvar, comp, logx=FALSE, outfile){

    jpeg(width=8, height=8, units='in', filename=outfile, type='cairo',
         res=300, quality=100)
    # pdf(outfile, height=8, width=8)
    par(mar=c(5, 5, 4, 6))

    plotcol = case_when(comp == 'GPP_site_mean' ~ gppcolor,
                        comp == 'ER_site_mean' ~ ercolor,
                        comp == 'NEP_site_mean' ~ 'black')

    if(xvar == 'Stream_PAR_sum'){
        xxlab = 'Light Availability (Mean Annual Surface PAR)'
    } else {
        xxlab = 'MODIS NPP'
    }

    gpprng = range(sp_site$GPP_site_mean, na.rm=TRUE)
    if(comp == 'ER_site_mean'){
        xx = abs(sp_site[[comp]])
    } else {
        xx = sp_site[[comp]]
    }
    if(comp != 'NEP_site_mean'){
        rescaled = ((xx - gpprng[1]) /
                        (gpprng[2] - gpprng[1])) * (5 - 1) + 1
    } else {
        xxrng = range(xx, na.rm=TRUE)
        # rescaled = ((xx - xxrng[1]) /
        #                 (xxrng[2] - xxrng[1])) * (3) + 0.2
        rescaled = ((xx - -2000) /
                        (1000 - -2000)) * (4) + 0.5
    }

    xxvar = sp_site[[xvar]]
    if(logx){
        xxvar = log(xxvar)
        if(xvar == 'Stream_PAR_sum'){
            xlm = c(.5, 2.8)
        } else {
            xlm = c(3.95, 7.2)
        }
    } else {
        if(xvar == 'Stream_PAR_sum'){
            xlm = c(2, 15.5)
        } else {
            xlm = range(xxvar, na.rm=TRUE)
        }
    }

    plot(xxvar,
         sp_site$Disch_ar1, pch=21,
    # plot(xxvar[sp_high_cov_bool],
    #      sp_site$Disch_ar1[sp_high_cov_bool], pch=21,
         xlab=xxlab, xaxt='n',
         ylab='Predictability of Flow (Discharge AR-1 Coeff.)',
         col=alpha(plotcol, alpha=0.5), bty='o',
         ylim=c(0.1, 1), xlim=xlm,
         xpd=NA, main='', cex=rescaled, font.lab=2,
         # xpd=NA, main='', cex=rescaled[sp_high_cov_bool], font.lab=2,
         bg=alpha(plotcol, alpha=0.5))
    # points(xxvar[! sp_high_cov_bool],
    #        sp_site$Disch_ar1[! sp_high_cov_bool], pch=21,
    #        col=alpha(plotcol, alpha=0.5), bg='transparent',
    #        cex=rescaled[! sp_high_cov_bool])

    if(comp != 'NEP_site_mean'){
        legend('right', legend=c(expression(paste(0.01)), '', '',
                                 expression(paste(4000))),
               bty='n', pch=21, pt.bg='transparent', x.intersp=1.7,
               pt.cex=c(1, 2, 3, 5), col='gray30', xpd=NA,
               y.intersp=c(1, 2, 1.2, 1.6),
               inset=c(-0.16, 0), title='')
    } else {
        legend('right', legend=c(expression(paste(-2000)), '', '',
                                 expression(paste(1000))),
               bty='n', pch=21, pt.bg='transparent', x.intersp=1.7,
               pt.cex=seq(0.5, 4, length.out=4), col='gray30', xpd=NA,
               y.intersp=c(1, 2, 1.2, 1.6),
               inset=c(-0.18, 0), title='')
    }
    if(comp == 'GPP_site_mean'){
        legend('right', legend=c('','','',''),
               bty='n', pch=21, pt.bg='transparent', x.intersp=1.5,
               pt.cex=c(1, 2, 3, 5), col='transparent', xpd=NA,
               inset=c(-0.15, 0),
               title=expression(paste(bold('Cumul.\nAnnual\nGPP'))))
    } else if(comp == 'ER_site_mean'){
        legend('right', legend=c('','','',''),
               bty='n', pch=21, pt.bg='transparent', x.intersp=1.5,
               pt.cex=c(1, 2, 3, 5), col='transparent', xpd=NA,
               inset=c(-0.15, 0),
               title=expression(paste(bold('Cumul.\nAnnual\nER'))))
    } else {
        legend('right', legend=c('','','',''),
               bty='n', pch=21, pt.bg='transparent', x.intersp=1.5,
               pt.cex=c(1, 2, 3, 5), col='transparent', xpd=NA,
               inset=c(-0.15, 0),
               title=expression(paste(bold('Cumul.\nAnnual\nNEP'))))
    }

    if(logx){
        if(xvar == 'Stream_PAR_sum'){
            tcks = c(1, 2, 4, 8, 16)
        } else {
            tcks = c(200, 400, 800, 1600, 3200)
        }
        tcks_log = log(tcks)
        axis(1, at=tcks_log, labels=tcks)
    } else {
        axis(1)
    }

    dev.off()
}

bubble_plot(xvar='Stream_PAR_sum', comp='GPP_site_mean', logx=FALSE,
            outfile=here('figures', 'bubble_plots', 'PAR_GPP_linear.jpeg'))
bubble_plot(xvar='Stream_PAR_sum', comp='GPP_site_mean', logx=TRUE,
            outfile=here('figures', 'bubble_plots', 'PAR_GPP_log.jpeg'))
bubble_plot(xvar='Stream_PAR_sum', comp='ER_site_mean', logx=FALSE,
            outfile=here('figures', 'bubble_plots', 'PAR_ER_linear.jpeg'))
bubble_plot(xvar='Stream_PAR_sum', comp='ER_site_mean', logx=TRUE,
            outfile=here('figures', 'bubble_plots', 'PAR_ER_log.jpeg'))
bubble_plot(xvar='Stream_PAR_sum', comp='NEP_site_mean', logx=FALSE,
            outfile=here('figures', 'bubble_plots', 'PAR_NEP_linear.jpeg'))
bubble_plot(xvar='Stream_PAR_sum', comp='NEP_site_mean', logx=TRUE,
            outfile=here('figures', 'bubble_plots', 'PAR_NEP_log.jpeg'))
bubble_plot(xvar='MOD_ann_NPP', comp='GPP_site_mean', logx=FALSE,
            outfile=here('figures', 'bubble_plots', 'MODNPP_GPP_linear.jpeg'))
bubble_plot(xvar='MOD_ann_NPP', comp='GPP_site_mean', logx=TRUE,
            outfile=here('figures', 'bubble_plots', 'MODNPP_GPP_log.jpeg'))
bubble_plot(xvar='MOD_ann_NPP', comp='ER_site_mean', logx=FALSE,
            outfile=here('figures', 'bubble_plots', 'MODNPP_ER_linear.jpeg'))
bubble_plot(xvar='MOD_ann_NPP', comp='ER_site_mean', logx=TRUE,
            outfile=here('figures', 'bubble_plots', 'MODNPP_ER_log.jpeg'))
bubble_plot(xvar='MOD_ann_NPP', comp='NEP_site_mean', logx=FALSE,
            outfile=here('figures', 'bubble_plots', 'MODNPP_NEP_linear.jpeg'))
bubble_plot(xvar='MOD_ann_NPP', comp='NEP_site_mean', logx=TRUE,
            outfile=here('figures', 'bubble_plots', 'MODNPP_NEP_log.jpeg'))
```

# Figure S1: sitemap

See code/paper for details

```{r Figure S1, eval = FALSE, include = FALSE}

library(sf)
library(tmap)
library(tmaptools)
library(OpenStreetMap) #requires rJava package (which requires local JRE, JDK installation)

sites_fnet = readRDS(here('output_data', 'fluxnet_site_info_filtered.rds')) %>%
    as_tibble() %>%
    select(site_code = Site_ID, Lat, Lon) %>%
    mutate(source = 'FLUXNET') %>%
    filter(! is.na(Lon)) %>%
    st_as_sf(coords = c('Lon', 'Lat'),
             remove = FALSE,
             crs = 4326)
    
sites_sp = readRDS(here('output_data', 'lotic_site_info_filtered.rds')) %>%
    as_tibble() %>%
    select(site_code = Site_ID, Lat, Lon) %>%
    mutate(source = 'StreamPULSE') %>%
    st_as_sf(coords = c('Lon', 'Lat'),
             remove = FALSE,
             crs = 4326)

basemap = tmaptools::read_osm(sites_fnet,
                              # ext = 1.1,
                              relative = TRUE,
                              ylim = c(-0.04, 1.01),
                              xlim = c(-0.04, 1.02),
                              type = 'nps')

tmap_mode('plot')

map = tm_shape(basemap) + tm_rgb() +
    tm_shape(sites_sp) + tm_symbols(shape = 21, col = 'steelblue3', size = 0.1,
                                    border.lwd = 0.5, border.col = 'black') +
    tm_shape(sites_fnet) + tm_symbols(shape = 21, col = fnetcolor, size = 0.1,
                                      border.lwd = 0.5, border.col = 'black') +
    # tm_compass(type = 'arrow', position = c(0.1, 0.01), show.labels = 1,
    #            size = 2, text.size = 1) +
    # tm_style(style='white') +
    # tm_layout(frame=TRUE, bg.color='white') +
    tm_scale_bar(text.size = 0.6, position=c(0, 0), width = 0.2) +
    tm_add_legend(type = 'symbol', labels = 'StreamPULSE', col = 'steelblue3',
                  shape = 21, border.col = 'black', border.lwd = 1) +
    tm_add_legend(type = 'symbol', labels = 'FLUXNET', col = fnetcolor,
                  shape = 21, border.col = 'black', border.lwd = 1) +
    tm_legend(show = TRUE, position = c(0.005, 0.09), outside = FALSE,
              bg.color = 'gray94', frame = TRUE, text.size = 0.6,
              between.margin = 0)

tmap_save(map, filename = here('figures', 'sitemap.jpg'), bg = 'white',
          dpi = 800, width = 6, units = 'in')
          # compression = 'lzw',
```

# Figure S2: peak months

See code/paper for details

```{r Figure Y, include = FALSE}
dir.create(here('figures', 'peak_months'), showWarnings = FALSE)

library(lubridate)

error.bars <- function(x, y, upper, lower=upper, cap.length=0.1, horiz=F,...){
    if(length(x) != length(y) | length(y) !=length(lower) | length(lower) != length(upper))
        stop("One or more vectors is not the same length")

    if(horiz==F) {
        arrows(x,y+upper, x, y-lower, angle=90, code=3, length=cap.length, ...)
    } else if (horiz==T) {
        arrows(x+upper,y, x-lower, y, angle=90, code=3, length=cap.length, ...)
    }
}

sitemonths_sp = sp_full %>%
    rename(GPP = GPP_C_filled,
           ER = ER_C_filled) %>%
    mutate(month = month(Date)) %>%
    group_by(sitecode, Year, month) %>%
    summarize(across(c(GPP, ER), mean, na.rm = TRUE),
              .groups = 'drop') %>%
    group_by(sitecode, month) %>%
    summarize(across(c(GPP, ER), #mean, na.rm = TRUE),
                     list(mean = ~mean(., na.rm = TRUE),
                          sd = ~sd(., na.rm = TRUE))),
              n = n(),
              .groups = 'drop')

sitemonths_fnet = fnet_full %>%
    mutate(month = month(Date)) %>%
    group_by(sitecode, Year, month) %>%
    summarize(across(c(GPP, ER), mean, na.rm = TRUE),
              .groups = 'drop') %>%
    group_by(sitecode, month) %>%
    summarize(across(c(GPP, ER), #mean, na.rm = TRUE),
                     list(mean = ~mean(., na.rm = TRUE),
                          sd = ~sd(., na.rm = TRUE))),
              n = n(),
              .groups = 'drop')

#GPP

maxmonths_etc_sp_gpp = sitemonths_sp %>%
    group_by(sitecode) %>%
    filter(GPP_mean == max(GPP_mean),
           n == max(n)) %>%
    summarize(GPP_mean = first(GPP_mean),
              GPP_sd = first(GPP_sd),
              n = first(n),
              month = first(month),
              .groups = 'drop') %>%
    mutate(SE = GPP_sd / sqrt(n))

maxmonths_etc_fnet_gpp = sitemonths_fnet %>%
    group_by(sitecode) %>%
    filter(GPP_mean == max(GPP_mean),
           n == max(n)) %>%
    summarize(GPP_mean = first(GPP_mean),
              GPP_sd = first(GPP_sd),
              n = first(n),
              month = first(month),
              .groups = 'drop') %>%
    mutate(SE = GPP_sd / sqrt(n))

maxmonths_sp_gpp = c('1' = 0, table(maxmonths_etc_sp_gpp$month), '12' = 0)
maxmonths_fnet_gpp = table(maxmonths_etc_fnet_gpp$month)
maxmonths_gpp = matrix(c(maxmonths_fnet_gpp, maxmonths_sp_gpp),
                   nrow = 2, byrow = TRUE,
                   dimnames = list(c('FLUXNET', 'StreamPULSE'),
                                   month.abb))

jpeg(width=8, height=8, units='in', res=300, quality=100, type='cairo',
     filename=here('figures', 'peak_months/GPP.jpg'))

barplot(maxmonths_gpp, border = NA, beside = TRUE,
        col = c(fnetcolor, spcolor),
        ylab = 'n', main = 'Highest GPP months')
legend(x=25, y=60, legend=c('FLUXNET', 'StreamPULSE'),
       fill = c(fnetcolor, spcolor), bty = 'n', border = NA)
text(x=25, y=52, adj = 0, paste('n FLUXNET =', sum(maxmonths_fnet_gpp)))
text(x=25, y=49, adj = 0, paste('n StreamPULSE =', sum(maxmonths_sp_gpp)))

dev.off()

#ER

maxmonths_etc_sp_er = sitemonths_sp %>%
    group_by(sitecode) %>%
    filter(ER_mean == min(ER_mean),
           n == max(n)) %>%
    summarize(ER_mean = first(ER_mean),
              ER_sd = first(ER_sd),
              n = first(n),
              month = first(month),
              .groups = 'drop') %>%
    mutate(SE = ER_sd / sqrt(n))

maxmonths_etc_fnet_er = sitemonths_fnet %>%
    group_by(sitecode) %>%
    filter(ER_mean == min(ER_mean),
           n == max(n)) %>%
    summarize(ER_mean = first(ER_mean),
              ER_sd = first(ER_sd),
              n = first(n),
              month = first(month),
              .groups = 'drop') %>%
    mutate(SE = ER_sd / sqrt(n))

maxmonths_sp_er = table(maxmonths_etc_sp_er$month)
maxmonths_fnet_er = table(maxmonths_etc_fnet_er$month)
# maxmonths_fnet_er = c(tt[1:9], c('10' = 0), tt[11:12])
maxmonths_er = matrix(c(maxmonths_fnet_er, maxmonths_sp_er),
                   nrow = 2, byrow = TRUE,
                   dimnames = list(c('FLUXNET', 'StreamPULSE'),
                                   month.abb))

jpeg(width=8, height=8, units='in', res=300, quality=100, type='cairo',
     filename=here('figures', 'peak_months/ER.jpg'))

barplot(maxmonths_er, border = NA, beside = TRUE,
        col = c(fnetcolor, spcolor),
        ylab = 'n', main = 'Highest ER months')
legend(x=25, y=55, legend=c('FLUXNET', 'StreamPULSE'),
       fill = c(fnetcolor, spcolor), bty = 'n', border = NA)
text(x=25, y=48, adj = 0, paste('n FLUXNET =', sum(maxmonths_fnet_er)))
text(x=25, y=45, adj = 0, paste('n StreamPULSE =', sum(maxmonths_sp_er)))

dev.off()

```
